Share Buyback and Equity Issue Anomalies Revisited: Interactive Tool

### 2016-02-24 bb_issuers.Rnw
tool

%%Blind title page
%\begin{titlepage}
%\begin{center}
%  \centering
%\vskip 60pt
%\LARGE Share Buyback and Equity Issue Anomalies Revisited \par
%\vskip 1.5em
%\normalsize  \today
%\noindent \\ 
%\begin{abstract}
%\noindent We re-examine the behavior of stock returns after share buyback and equity issuance announcements using the five-factor model proposed by \cite{FamaFrench2015a}. We confirm the findings of \cite{FamaFrench2015b} that the equity issue anomaly, i.e., the fact that equity issues are followed by negative long-term excess returns, disappears after replacing the \cite{FamaFrench1993} three-factor model with the five-factor model. However, long term positive excess returns after buyback announcements, first reported by \cite{IkenberryLakonishokVermaelen1995}, remain economically and statistically significant. Moreover, the Undervaluation Index proposed by \cite{PeyerVermaelen2009} remains a useful proxy for the likelihood that the buyback is driven by undervaluation. The buyback anomaly is robust over time and across sectors. Firms with low correlations with the market seem to be better at market timing, which is consistent with the hypothesis that the repurchase is driven by superior company-specific information. Firms with high pre-announce stock returns volatility also have higher abnormal returns, which is consistent with the hypothesis that a repurchase authorization is an option to take advantage of undervaluation as argued by \cite{IkenberryVermaelen1996}. This result is also consistent with \cite{Stambaugh2015} who find a positive relation between stock returns and idiosyncratic volatility for undervalued stocks, unlike the well-known low volatility anomaly across all stocks. Finally, based on these findings, we develop an enhanced Undervaluation Index that improves abnormal returns relative to the index developed by \cite{PeyerVermaelen2009}.
%\end{abstract}
%\vspace{2cm}
%\end{center}
%\medskip
%\noindent \textit{JEL classification}: G35
%\medskip
%\noindent \textit{Keywords}: Share Buybacks, Seasoned Equity Offerings
%\end{titlepage}

@@

\vspace{-.2in}
\begin{abstract}
\noindent We re-examine the behavior of stock returns after share buyback and equity issuance announcements using the five-factor model proposed by \cite{FamaFrench2015a}. We confirm the findings of \cite{FamaFrench2015b} that the equity issue anomaly, i.e., the fact that equity issues are followed by negative long-term excess returns, disappears after replacing the \cite{FamaFrench1993} three-factor model with the five-factor model. However, long term positive excess returns after buyback announcements, first reported by \cite{IkenberryLakonishokVermaelen1995}, remain economically and statistically significant. Moreover, the Undervaluation Index proposed by \cite{PeyerVermaelen2009} remains a useful proxy for the likelihood that the buyback is driven by undervaluation. The buyback anomaly is robust over time and across sectors. Firms with low correlations with the market seem to be better at market timing, which is consistent with the hypothesis that the repurchase is driven by superior company-specific information. Firms with high pre-announce stock returns volatility also have higher abnormal returns, which is consistent with the hypothesis that a repurchase authorization is an option to take advantage of undervaluation as argued by \cite{IkenberryVermaelen1996}. This result is also consistent with \cite{Stambaugh2015} who find a positive relation between stock returns and idiosyncratic volatility for undervalued stocks, unlike the well-known low volatility anomaly across all stocks. Finally, based on these findings, we develop an enhanced Undervaluation Index that improves abnormal returns relative to the index developed by \cite{PeyerVermaelen2009}.

>>>>

%%Blind title page
%\begin{titlepage}
%\begin{center}
%  \centering
%\vskip 60pt
%\LARGE Share Buyback and Equity Issue Anomalies Revisited \par
%\vskip 1.5em
%\normalsize  \today
%\noindent \\ 
%\begin{abstract}
%\noindent We re-examine the behavior of stock returns after net issue  announcements using the five-factor model proposed by Fama and French (2015a). We confirm their findings that the equity issue anomaly disappears  with the five-factor model, but the buyback anomaly does not.. Buyback announcements by firms  with high pre-announcement (idiosyncratic) volatility  are followed by higher long-term abnormal returns. This is inconsistent with the widely documented low volatility anomaly, but consistent with Stambaugh, Yu, and Yuan (2015) who find a positive relation between stock returns and idiosyncratic volatility for undervalued stocks.
%\end{abstract}
%\vspace{2cm}
%\end{center}
%\medskip
%\noindent \textit{JEL classification}: G35
%\medskip
%\noindent \textit{Keywords}: Share Buybacks, Seasoned Equity Offerings
%\end{titlepage}

@@

\vspace{-.2in}
\begin{abstract}
\noindent We re-examine the behavior of stock returns after net issue  announcements using the five-factor model proposed by Fama and French (2015a). We confirm their findings that the equity issue anomaly disappears  with the five-factor model, but the buyback anomaly does not. Buyback announcements by firms  with high pre-announcement (idiosyncratic) volatility  are followed by higher long-term abnormal returns. This is inconsistent with the widely documented low volatility anomaly, but consistent with Stambaugh, Yu, and Yuan (2015) who find a positive relation between stock returns and idiosyncratic volatility for undervalued stocks.

### 2016-02-24 index.Rmd
tool

---
title: "Share Buyback and Equity Issue Anomalies Revisited: Interactive Tool"
author: "[Theodoros Evgeniou](http://faculty.insead.edu/theodoros-evgeniou/), [Enric Junque de Fortuny](http://ciri.be/), [Nick Nassuphis](https://www.linkedin.com/in/nick-nassuphis-8029372), and [Theo Vermaelen](http://faculty.insead.edu/theo-vermaelen/)"
output:  
 html_document:
 fig_width: 6
 fig_height: 2.5
 fig_align: center
 in_header: header.html
 runtime: shiny
runtime: shiny
---

@@

#if(1) {
#  CACHE <- car_table_cache(DATASET_INI$returns_by_event_monthly,
#                           DATASET_INI$SDC$Event.Date,Risk_Factors_Monthly,formula_used = five_factor_model)
#  save("CACHE",file = "cache.Rdata")
#} else {
#  load("cache.Rdata")  
#}
report_months_cal = c(as.numeric(reported_times[1:(length(reported_times)-1)]),1)

@@

```{r, fig.width=6, fig.height=5,echo=FALSE}
downloadHandler(
  filename = "report.pdf",
  content = function(file) {
    cat("knitting")
    #knitr::knit("report.Rnw",envir=environment())
    cat("texing")
    #tools::texi2pdf("report.tex")
    knitr::knit2pdf("report.Rnw",envir=globalenv())
    #toremove <- c("report.aux","report.bbl","report.blg","report.log","report.out")
    #file.remove(normalizePath(toremove))
    cat("done")
    file.rename(normalizePath('report.pdf'), file)
  }
)
```

@@

```{r, eval=TRUE, echo = FALSE}
acceptaple_features = c("Prior.Returns.Score", "BE.ME.Score","Size.Score" ,"Valuation.Index","Industry", "Market.Cap",          "Percent.Shares", "Purpose.Code","Source.of.funds","Market.beta","SMB.beta","HML.beta" , "RMW.beta" ,"CMA.beta", "one.minus.Rsq.Score", "pre.vol.Score","ivol", "leverage.ratio")

@@

```{r, eval=TRUE, echo = FALSE}
acceptaple_features1 = c("Prior.Returns.Score", "BE.ME.Score","Size.Score" ,"Valuation.Index", "Market.Cap",          "Percent.Shares", "Market.beta","SMB.beta","HML.beta" , "RMW.beta" ,"CMA.beta", "one.minus.Rsq.Score", "pre.vol.Score","ivol", "leverage.ratio")

@@

feature_IRATStable <- reactive({
  thesign = ifelse(input$characteristic_irats_sign, -1, +1)
  thefeature = thesign*DATASET_NEW()$SDC[[which(names(DATASET_NEW()$SDC) == input$characteristic_irats)]]  
  High_feature_events = which(scrub(thefeature) >= quantile(thefeature[!is.na(thefeature)],1-as.numeric(input$quantile_used)) & !is.na(thefeature))
  Low_feature_events = which(scrub(thefeature) <= quantile(thefeature[!is.na(thefeature)],as.numeric(input$quantile_used)) & !is.na(thefeature))
  res = round(cbind(
    #car_table(DATASET_NEW()$returns_by_event_monthly[,Low_feature_events], DATASET_NEW()$SDC$Event.Date[Low_feature_events], Risk_Factors_Monthly)$results,
    #car_table(DATASET_NEW()$returns_by_event_monthly[,High_feature_events], DATASET_NEW()$SDC$Event.Date[High_feature_events], Risk_Factors_Monthly)$results
    car_table_cached(DATASET_NEW()$CACHE,Low_feature_events)$results,
    car_table_cached(DATASET_NEW()$CACHE,High_feature_events)$results
  ),2)[reported_times,]
  colnames(res) <- c("Low: CAR", "t-stat","p-value", "High: CAR", "t-stat","p-value")
  as.data.frame(res)
})

feature_CALtable <- reactive({
  thesign = ifelse(input$characteristic_irats_sign, -1, +1)
  thefeature = thesign*DATASET_NEW()$SDC[[which(names(DATASET_NEW()$SDC) == input$characteristic_irats)]]  
  High_feature_events = which(scrub(thefeature) >= quantile(thefeature[!is.na(thefeature)],1-as.numeric(input$quantile_used)) & !is.na(thefeature))
  Low_feature_events = which(scrub(thefeature) <= quantile(thefeature[!is.na(thefeature)],as.numeric(input$quantile_used)) & !is.na(thefeature))
  res = round(cbind(
    calendar_table2(DATASET_NEW()$returns_by_event_monthly[,Low_feature_events], DATASET_NEW()$SDC$Event.Date[Low_feature_events], Risk_Factors_Monthly,report_months_cal)$results,
    calendar_table2(DATASET_NEW()$returns_by_event_monthly[,High_feature_events], DATASET_NEW()$SDC$Event.Date[High_feature_events], Risk_Factors_Monthly,report_months_cal)$results
  ),2)[reported_times,]
  colnames(res) <- c("Low: CAL", "t-stat","p-value", "High: CAL", "t-stat","p-value")
  as.data.frame(res)
})

High_feature_Hedged_react <- reactive({
  thefeature = DATASET_NEW()$SDC[[which(names(DATASET_NEW()$SDC) == input$characteristic_irats)]]  
  High_feature_events = which(scrub(thefeature) >= quantile(thefeature[!is.na(thefeature)],1-as.numeric(input$quantile_used)) & !is.na(thefeature))
  High_feature <- apply(PNL_matrix_BB(start_date_event,"One.Year.After", High_feature_events,  DATASET_NEW()$DatesMonth, DATASET_NEW()$returns_by_event_monthly,event=1),1,non_zero_mean)
  remove_initialization_time(suppressWarnings(scrub(alpha_lm(High_feature,Risk_Factors_Monthly[,pnl_hedge_factors],hedge_months,trade=1))),min_date=FirstTrade)
})

Low_feature_Hedged_react <- reactive({
  thefeature = DATASET_NEW()$SDC[[which(names(DATASET_NEW()$SDC) == input$characteristic_irats)]]  
  Low_feature_events = which(scrub(thefeature) <= quantile(thefeature[!is.na(thefeature)],as.numeric(input$quantile_used)) & !is.na(thefeature))
  Low_feature <- apply(PNL_matrix_BB(start_date_event,"One.Year.After", Low_feature_events,  DATASET_NEW()$DatesMonth, DATASET_NEW()$returns_by_event_monthly,event=1),1,non_zero_mean)
  remove_initialization_time(suppressWarnings(scrub(alpha_lm(Low_feature,Risk_Factors_Monthly[,pnl_hedge_factors],hedge_months,trade=1))),min_date=FirstTrade)  
})

@@

Much like the EU-index in the paper, one can create other indices by combining different firm characteristics as needed.

```{r, eval=TRUE, echo = FALSE}
acceptaple_features2 = c("Prior.Returns.Score", "BE.ME.Score","Size.Score" ,"Valuation.Index", "Market.Cap",          "Percent.Shares", "Market.beta","SMB.beta","HML.beta" , "RMW.beta" ,"CMA.beta", "one.minus.Rsq.Score", "pre.vol.Score","ivol", "leverage.ratio","None")
```

<br>
```{r, eval=TRUE, echo = FALSE}
selectInput("characteristic_irats1", "Select First Firm Characteristic for your new index:", choices = acceptaple_features2,selected = acceptaple_features2[1], width = "1000px")
checkboxInput("characteristic_irats1_sign", "If you would like to change the sign of this firm characteristic (Low will be High and vice versa), select here:", value = FALSE, width = "1000px")
```

<br>
```{r, eval=TRUE, echo = FALSE}
selectInput("characteristic_irats2", "Select Second Firm Characteristic for your new index:", choices = acceptaple_features2,selected ="None", width = "1000px")
checkboxInput("characteristic_irats2_sign", "If you would like to change the sign of this firm characteristic (Low will be High and vice versa), select here:", value = FALSE, width = "1000px")
```

<br>
```{r, eval=TRUE, echo = FALSE}
selectInput("characteristic_irats3", "Select Third Firm Characteristic for your new index:", choices = acceptaple_features2,selected = "None", width = "1000px")
checkboxInput("characteristic_irats3_sign", "If you would like to change the sign of this firm characteristic (Low will be High and vice versa), select here:", value = FALSE, width = "1000px")
```

<br>
```{r, eval=TRUE, echo = FALSE}
selectInput("characteristic_irats4", "Select Fourth Firm Characteristic for your new index:", choices = acceptaple_features2,selected = "None", width = "1000px")
checkboxInput("characteristic_irats4_sign", "If you would like to change the sign of this firm characteristic (Low will be High and vice versa), select here:", value = FALSE, width = "1000px")
```

<br>
```{r, eval=TRUE, echo = FALSE}
selectInput("characteristic_irats5", "Select Fifth Firm Characteristic for your new index:", choices = acceptaple_features2,selected = "None", width = "1000px")
checkboxInput("characteristic_irats5_sign", "If you would like to change the sign of this firm characteristic (Low will be High and vice versa), select here:", value = FALSE, width = "1000px")
```

<br>
```{r, eval=TRUE, echo = FALSE}
selectInput("characteristic_irats6", "Select Sixth Firm Characteristic for your new index:", choices = acceptaple_features2,selected = "None", width = "1000px")
checkboxInput("characteristic_irats6_sign", "If you would like to change the sign of this firm characteristic (Low will be High and vice versa), select here:", value = FALSE, width = "1000px")
```

>>>>

---
title: "Share Buyback and Equity Issue Anomalies Revisited: Interactive Tool"
author: "[Theodoros Evgeniou](http://faculty.insead.edu/theodoros-evgeniou/), [Enric Junqué de Fortuny](http://ciri.be/), [Nick Nassuphis](https://www.linkedin.com/in/nick-nassuphis-8029372), and [Theo Vermaelen](http://faculty.insead.edu/theo-vermaelen/)"
output:  
 html_document:
 fig_width: 6
 fig_height: 2.5
 fig_align: center
 in_header: header.html
 runtime: shiny
runtime: shiny
---

@@

#if(1) {
#  CACHE <- car_table_cache(DATASET_INI$returns_by_event_monthly,
#                           DATASET_INI$SDC$Event.Date,Risk_Factors_Monthly,formula_used = five_factor_model)
#  save("CACHE",file = "cache.Rdata")
#} else {
#  load("cache.Rdata")  
#}
report_months_cal = c(as.numeric(reported_times[1:(length(reported_times)-1)]),1)
```

@@

```{r, fig.width=6, fig.height=5,echo=FALSE}
downloadHandler(
  filename = "report.pdf",
  content = function(file) {
    knitr::knit2pdf("report.Rnw",envir=environment())
    file.rename(normalizePath('report.pdf'), file)
  }
)
```

@@

```{r, eval=TRUE, echo = FALSE}
acceptaple_features = c("Prior.Returns.Score", "BE.ME.Score","Size.Score" ,"Valuation.Index", "Market.Cap",          "Percent.Shares","Market.beta","SMB.beta","HML.beta" , "RMW.beta" ,"CMA.beta", "one.minus.Rsq.Score", "pre.vol.Score","ivol", "leverage.ratio","Idiosyncraticness","Industry", "Purpose.Code","Source.of.funds")

@@

```{r, eval=TRUE, echo = FALSE}
acceptaple_features1 = c("Prior.Returns.Score", "BE.ME.Score","Size.Score" ,"Valuation.Index", "Market.Cap",          "Percent.Shares", "Market.beta","SMB.beta","HML.beta" , "RMW.beta" ,"CMA.beta", "one.minus.Rsq.Score", "pre.vol.Score","ivol", "leverage.ratio","Idiosyncraticness")

@@

feature_IRATStable <- reactive({
  thesign = ifelse(input$characteristic_irats_sign, -1, +1)
  thefeature = thesign*DATASET_NEW()$SDC[[which(names(DATASET_NEW()$SDC) == input$characteristic_irats)]]

  thefeature = as.numeric(scrub(thefeature))
  quantile_used_ind = as.numeric(input$quantile_used_ind)

  High_feature_events = !is.na(thefeature) & scrub(thefeature) >= quantile(scrub(thefeature[!is.na(thefeature)]),1-quantile_used_ind)
  Low_feature_events  = !is.na(thefeature) & scrub(thefeature) <= quantile(scrub(thefeature[!is.na(thefeature)]),quantile_used_ind)

  #High_feature_events = which(scrub(thefeature) > quantile(thefeature[!is.na(thefeature)],1-as.numeric(input$quantile_used)) & !is.na(thefeature))
  #Low_feature_events = which(scrub(thefeature) < quantile(thefeature[!is.na(thefeature)],as.numeric(input$quantile_used)) & !is.na(thefeature))
  res = round(cbind(
    #car_table(DATASET_NEW()$returns_by_event_monthly[,Low_feature_events], DATASET_NEW()$SDC$Event.Date[Low_feature_events], Risk_Factors_Monthly)$results,
    #car_table(DATASET_NEW()$returns_by_event_monthly[,High_feature_events], DATASET_NEW()$SDC$Event.Date[High_feature_events], Risk_Factors_Monthly)$results
    car_table_cached(DATASET_NEW()$CACHE,Low_feature_events)$results,
    car_table_cached(DATASET_NEW()$CACHE,High_feature_events)$results
  ),2)[reported_times,]
  colnames(res) <- c("Low: CAR", "t-stat","p-value", "High: CAR", "t-stat","p-value")
  as.data.frame(res)
})

feature_CALtable <- reactive({
  thesign = ifelse(input$characteristic_irats_sign, -1, +1)
  thefeature = thesign*DATASET_NEW()$SDC[[which(names(DATASET_NEW()$SDC) == input$characteristic_irats)]]  
  High_feature_events = which(scrub(thefeature) > quantile(thefeature[!is.na(thefeature)],1-as.numeric(input$quantile_used)) & !is.na(thefeature))
  Low_feature_events = which(scrub(thefeature) < quantile(thefeature[!is.na(thefeature)],as.numeric(input$quantile_used)) & !is.na(thefeature))
  res = round(cbind(
    calendar_table2(DATASET_NEW()$returns_by_event_monthly[,Low_feature_events], DATASET_NEW()$SDC$Event.Date[Low_feature_events], Risk_Factors_Monthly,report_months_cal)$results,
    calendar_table2(DATASET_NEW()$returns_by_event_monthly[,High_feature_events], DATASET_NEW()$SDC$Event.Date[High_feature_events], Risk_Factors_Monthly,report_months_cal)$results
  ),2)[reported_times,]
  colnames(res) <- c("Low: CAL", "t-stat","p-value", "High: CAL", "t-stat","p-value")
  as.data.frame(res)
})

High_feature_Hedged_react <- reactive({
  thefeature = as.numeric(DATASET_NEW()$SDC[[which(names(DATASET_NEW()$SDC) == input$characteristic_irats)]])
  quantile_used <- as.numeric(input$quantile_used)

  High_feature_events = !is.na(thefeature) & scrub(thefeature) >= quantile(scrub(thefeature[!is.na(thefeature)]),1-quantile_used)

  High_feature <- apply(PNL_matrix_BB(start_date_event,"One.Year.After", High_feature_events,  DATASET_NEW()$DatesMonth, DATASET_NEW()$returns_by_event_monthly,event=1),1,non_zero_mean)
  remove_initialization_time(suppressWarnings(scrub(alpha_lm(High_feature,Risk_Factors_Monthly[,pnl_hedge_factors],hedge_months,trade=1))),min_date=FirstTrade)
})

Low_feature_Hedged_react <- reactive({
  thefeature = DATASET_NEW()$SDC[[which(names(DATASET_NEW()$SDC) == input$characteristic_irats)]]  
  Low_feature_events = which(scrub(thefeature) < quantile(thefeature[!is.na(thefeature)],as.numeric(input$quantile_used)) & !is.na(thefeature))
  Low_feature <- apply(PNL_matrix_BB(start_date_event,"One.Year.After", Low_feature_events,  DATASET_NEW()$DatesMonth, DATASET_NEW()$returns_by_event_monthly,event=1),1,non_zero_mean)
  remove_initialization_time(suppressWarnings(scrub(alpha_lm(Low_feature,Risk_Factors_Monthly[,pnl_hedge_factors],hedge_months,trade=1))),min_date=FirstTrade)  
})

@@

Much like the EU-index in the paper, one can create other indices by combining different firm characteristics as needed. You can select your own firm characteristics below and also indicate whether you would like to reverse the sign of this firm characteristic (Low will be High and vice versa).

```{r, eval=TRUE, echo = FALSE}
acceptaple_features2 = c("Prior.Returns.Score", "BE.ME.Score","Size.Score" ,"Valuation.Index", "Market.Cap",          "Percent.Shares", "Market.beta","SMB.beta","HML.beta" , "RMW.beta" ,"CMA.beta", "one.minus.Rsq.Score", "pre.vol.Score","ivol", "leverage.ratio","Idiosyncraticness","None")
```

<br>
```{r, eval=TRUE, echo = FALSE}
fluidPage( 
  h4("Firm Characteristics of the new index"),
  fluidRow(
    column(6,
      selectInput("characteristic_irats1", "First Firm Characteristic:", choices = acceptaple_features2,selected = acceptaple_features2[1], width = "425px"),
      checkboxInput("characteristic_irats1_sign", "Reverse", value = FALSE),
      selectInput("characteristic_irats2", "Second Firm Characteristic:", choices = acceptaple_features2,selected ="None", width = "425px"),
checkboxInput("characteristic_irats2_sign", "Reverse", value = FALSE),
    selectInput("characteristic_irats3", "Third Firm Characteristic:", choices = acceptaple_features2,selected = "None", width = "425px"),
checkboxInput("characteristic_irats3_sign", "Reverse", value = FALSE)
     ),
  column(6,
      selectInput("characteristic_irats4", "Fourth Firm Characteristic:", choices = acceptaple_features2,selected = "None", width = "425px"),
      checkboxInput("characteristic_irats4_sign", "Reverse", value = FALSE),
    selectInput("characteristic_irats5", "Fifth Firm Characteristic:", choices = acceptaple_features2,selected = "None", width = "425px"),
    checkboxInput("characteristic_irats5_sign", "Reverse", value = FALSE),
    selectInput("characteristic_irats6", "Sixth Firm Characteristic:", choices = acceptaple_features2,selected = "None", width = "425px"),
    checkboxInput("characteristic_irats6_sign", "Reverse", value = FALSE)
    )
  )
)
```
